|

### 2019-08-09 article/article.Rmd
edits intro, 1st par methods

Here we assess the impact of _Tá de Pé_ (TDP), a mobile phone application designed to lower the costs of evaluating public works and punishing political representatives in Brazil. Developed by Transparência Brasil[^transparencia], TDP allows Brazilian citizens to know the location of public school construction sites, check their completion status and anonymously request information from competent authorities. TDP users can also take pictures of the construction sites and submit them to volunteering engineers for examination. If the engineers classify the construction as delayed, TDP prompts users to send a message to the mayor's office asking for completion estimates and explanations about the construction status. TDP has been online since April 2017 in 1030 Brazilian municipalities and was the winner of the 2016 Google Social Impact grant with more than 200,000 popular votes[^google].

@@

We use the TDP app to conduct two experimental interventions and test its impact on six outcomes related to school completion rates and complaints to public authorities. Overall, our results show that providing information to citizens has very limited impact on policy outcomes. We find that the TDP app increased the likelihood of school completion by 6.8 percent in the first experiment, but the effect is null in the second intervention. We also find that pressure over the bureaucracy increased by 157.5 percent (SE = X.XX, t-test = X.XX) in our second experiment, although not in the first. Importantly, our two interventions have no significant impact on any of the four measures of school investment and completion rates we employed in this study. Community monitoring does not consistently change the behaviour of Brazilian politicians or civil servants.

Our results raise questions about the ability of citizens to hold representatives accountable using bottom-up strategies. @banerjee2018tangible, @bjorkman2009power and @fiala2017social find significant policy improvements after providing information to local communities, but our experiment is in line with @banerjee2010pitfalls, @keefer2014mass, @olken2007monitoring and @raffler2018weakness and offers little evidence in favour of information interventions. Our study suggests that other models of political accountability may be more effective for public service provision.

@@

Transparência Brasil partnered with the Brazilian branch of _Engineers without Borders_ (EWB), a non-governmental organisation[^ewb], to provide technical evaluations of school completion rates. The EWB bases their assessment on photos and GPS location sent through the app. The engineers' reports are later uploaded to the database and stored on the users' mobile devices so citizens can track the progress of the constructions.

@@

The

>>>>

Here we assess the impact of _Tá de Pé_ (TDP), a mobile phone application designed to lower the costs of evaluating public works and punishing political representatives in Brazil. Developed by Transparência Brasil[^transparencia], TDP allows Brazilian citizens to know the location of public school construction sites, check their completion status and anonymously request information from competent authorities. TDP users can also take pictures of the construction sites and submit them to independent engineers for examination. If the engineers classify the construction as delayed, TDP prompts users to send a message to the mayor's office asking for completion estimates and explanations about the construction status. TDP has been online since April 2017 in 1030 Brazilian municipalities and was the winner of the 2016 Google Social Impact grant with more than 200,000 popular votes[^google].

@@

We use the TDP app to conduct two experimental interventions and test its impact on six outcomes related to school completion rates and complaints to public authorities. Overall, our results show that providing information to citizens has very limited impact on policy outcomes. We find that the TDP app increased the likelihood of school completion by 6.8 percent in the first experiment, but the effect is null in the second intervention. We also find that pressure over the bureaucracy increased by 157.5 percent in our second experiment, although not in the first. Importantly, our two interventions have no significant impact on any of the four measures of school investment and completion rates we employed in this study.

The results raise questions about the ability of citizens to hold representatives accountable using bottom-up strategies. Although studies like @bjorkman2009power, @fiala2017social and @reinikka2005fighting report better policy outcomes after providing information to local communities, our findings are closer to @banerjee2010pitfalls, @keefer2014mass, @olken2007monitoring and @raffler2018weakness and offer mixed evidence that information interventions lead to greater government responsiveness. Our analysis suggests that grassroots monitoring does not affect government behaviour in Brazil, and other models of political accountability may be more effective for public service provision.

@@

Transparência Brasil partnered with the Brazilian branch of _Engineers without Borders_ (EWB), an independent non-governmental organisation[^ewb], to provide technical assessment of school completion rates. The EWB bases their evaluations on photos and GPS location sent through the app. The engineers' reports are later uploaded to the database and stored on the users' mobile devices so citizens can track the progress of the constructions.

@@

We implemented two interventions to measure the effect of the TDP app on school constructions outcomes. The first experiment ran from August 2017 to February 2018 and we carried out a follow-up study one year later, from August 2018 to February 2019.
